---
title: "Petrogale shape analysis"
author: "D. Rex Mitchell"
date: "2023-05-03"
output: html_document
editor_options: 
  chunk_output_type: console
---

IMPORT PACKAGES.
```{r}
library(geomorph)
library(landvR)
```

IMPORT SHAPE DATA.
```{r}
load("allometry_tests.rda")
```

PERFORM PROCRUSTES SUPERIMPOSITION AND REMOVE ASYMMETRY VARIATION.
```{r}
ind <- dimnames(alldata)[[3]]
pet_sym <- bilat.symmetry(alldata, ind=ind, land.pairs = sym, object.sym=T, curves=slide)
```

PERFORM STANDARD GPA TO OBTAIN CRANIAL (CENTROID) SIZES.
```{r}
gpa <- gpagen(alldata)
size <- gpa$Csize
```

TEST ALLOMETRY USING PROCRUSTES REGRESSION.
```{r}
allom <- procD.lm(pet_sym$symm.shape~log(size), iter=999); summary(allom)
allom_plot <- plot(allom, type = "regression", predictor = log(size), 
                   reg.type = "RegScore", pch = iconsPCA, bg=colsPCA, 
                   xlab = "log(centroid size)", ylab = "Regression score", cex=2)
```

PLOT ALLOMETRY WITH INTRASPECIFIC SLOPES.
```{r}
plot2 <- plot(allom_plot$RegScore~log(size), xlab = "log(cranial centroid size)", ylab="Shape regression score", method = "RegScore", pch = iconsPCA, bg=colsPCA, cex=2, main="shape~log(cranial centroid size")

col <- c("lightblue","red","red4","purple","red4","purple","purple4","blue","green","lightblue","purple4","orange","turquoise","yellow","lightblue","red","orange")

lm_temp <- NULL
xcoords <- vector()
ycoords <- vector ()

# this loop plots regression lines for each species
for (i in 1: length(levels (speciesnames))){
  xcoords <- subset(log(size), speciesnames == levels (speciesnames)[i] )
  ycoords <- subset(allom_plot$RegScore, speciesnames == levels (speciesnames)[i])
  lm_temp <- lm(ycoords~xcoords)
  segments(min(xcoords), lm_temp$coefficients[1] + lm_temp$coefficients[2] *min(xcoords), max(xcoords), lm_temp$coefficients[1] + lm_temp$coefficients[2] *max(xcoords), lwd = 4, col= col[i])
}
```

TEST ALLOMETRY INTERACTION.
```{r}
allomspec <- procD.lm(pet_sym$symm.shape~log(size)*speciesnames, iter=999); summary(allomspec)
allomspec_plot <- plotAllometry(allomspec, size = size, logsz = TRUE, 
                       method = "RegScore", pch = iconsPCA, bg=colsPCA, cex=2)
```

PLOT FULL MODEL WITH INTRASPECIFIC SLOPES.
```{R}
plot2 <- plot(allomspec_plot$RegScore~log(size), xlab = "log(cranial centroid size)", ylab="Shape regression score", method = "RegScore", pch = iconsPCA, bg=colsPCA, cex=2, main="shape~log(cranial centroid size)*species")

for (i in 1: length(levels (speciesnames))){
  xcoords <- subset(log(size), speciesnames == levels (speciesnames)[i] )
  ycoords <- subset(allomspec_plot$RegScore, speciesnames == levels (speciesnames)[i])
  lm_temp <- lm(ycoords~xcoords)
  segments(min(xcoords), lm_temp$coefficients[1] + lm_temp$coefficients[2] *min(xcoords), max(xcoords), lm_temp$coefficients[1] + lm_temp$coefficients[2] *max(xcoords), lwd = 4, col= col[i])
}
```

USE 'SHAPE.PREDICTOR' TO VISUALLY EXAMINE PREDICTED SHAPE CHANGES THAT OCCUR WITH INCREASED CRANIAL SIZE.
```{r}
preds <- shape.predictor(allom$GM$fitted, x = allom_plot$PredLine, 
                         predmin = min(allom_plot$PredLine), 
                         predmax = max(allom_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmax, preds$predmin, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)
```

COMPARE ALLOMETRIC TRAJECTORIES ACROSS SPECIES BY CALCULATING ANGLE VECTORS AND SIGNIFICANCE.
```{r}
allomspec <- procD.lm(pet_sym$symm.shape~log(size)*speciesnames, iter=999); summary(allomspec)
pw<-pairwise(allomspec,groups=speciesnames, covariate = log(size))
pw_sum <- summary(pw)
angle<-summary(pw,test="VC",angle.type="deg")
angle$summary.table
```

TO COMPARE INTRASPECIFIC ALLOMETRIC PREDICTIONS ACROSS SPECIES, USE THE FOLLOWING CODE AND SUBSTITUTE "SPECIES" WITH DESIRED SPECIFIC EPITHET (e.g., "Petrogale_SPECIES" --> "Petrogale_assimilis"). THIS PRODUCES PREDICTED SHAPE CHANGES FOR EACH SPECIES.
```{r}
speciesgp <- coords.subset(alldata, speciesnames)

ind <- dimnames(speciesgp$Petrogale_SPECIES)[[3]]
pet_sym <- bilat.symmetry(speciesgp$Petrogale_SPECIES, ind=ind, land.pairs = sym, object.sym=T, curves=slide)

gpa <- gpagen(speciesgp$Petrogale_SPECIES)
size <- gpa$Csize
range(size)

allom <- procD.lm(pet_sym$symm.shape~log(size), iter=999); summary(allom)
allom_plot <- plot(allom, type = "regression", predictor = log(size), 
                   reg.type = "RegScore", pch = 22, bg="lightblue", 
                   xlab = "log(centroid size)", ylab = "Regression score", cex=2)

preds <- shape.predictor(allom$GM$fitted, x = allom_plot$PredLine, 
                         predmin = min(allom_plot$PredLine), 
                         predmax = max(allom_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmax, preds$predmin, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 2, pt.size = 1)
```

PERFORM PRINCIPAL COMPONENT ANALYSIS AND PLOT PC1 VS PC2.
```{r}
PCA <- gm.prcomp(pet_sym$symm.shape)
plot(PCA$x[,2]~PCA$x[,1], pch=iconsPCA, cex=size/200, axes.linetype=NA, bg=colsPCA, asp=1, xlab = paste("PC1", "(", (signif(PCA$d[1]/sum(PCA$d)*100, digits=4)), "%)", sep=""), ylab = paste("PC2", "(", (signif(PCA$d[2]/sum(PCA$d)*100, digits=3)), "%)", sep=""))
```

TEST FOR SIZE CORRELATIONS FOR MAJOR COMPONENTS OF SHAPE.
```{r}
PC1cor <- lm(PCA$x[,1]~log(gpa$Csize)); print(summary(PC1cor))
PC2cor <- lm(PCA$x[,2]~log(gpa$Csize)); print(summary(PC2cor)) 
PC3cor <- lm(PCA$x[,3]~log(gpa$Csize)); print(summary(PC3cor)) 
PC4cor <- lm(PCA$x[,4]~log(gpa$Csize)); print(summary(PC4cor)) 
```

PLOT HEATMAPS OF SHAPE DIFFERENCES ALONG PC1.
```{r}
differences <- coordinates.difference(coordinates = PCA$shapes$shapes.comp1$max,
                                      reference = PCA$shapes$shapes.comp1$min, type = "spherical")
procrustes.var.plot(PCA$shapes$shapes.comp1$min, PCA$shapes$shapes.comp1$max, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)
```

PLOT HEATMAPS OF SHAPE DIFFERENCES ALONG PC2.
```{r}
differences <- coordinates.difference(coordinates = PCA$shapes$shapes.comp2$max,
                                      reference = PCA$shapes$shapes.comp2$min, type = "spherical")
procrustes.var.plot(PCA$shapes$shapes.comp2$min, PCA$shapes$shapes.comp2$max, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)
```

PLOT HEATMAPS OF SHAPE DIFFERENCES ALONG PC3.
```{r}
differences <- coordinates.difference(coordinates = PCA$shapes$shapes.comp3$max,
                                      reference = PCA$shapes$shapes.comp3$min, type = "spherical")
procrustes.var.plot(PCA$shapes$shapes.comp3$min, PCA$shapes$shapes.comp3$max, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)
```
