---
title: "Petrogale shape analysis"
author: "D. Rex Mitchell"
date: "2023-05-03"
output: html_document
---

IMPORT PACKAGES
```{r}
library(geomorph)
library(landvR)
```

IMPORT SHAPE DATA
```{r}
load("allometry_tests.rda")
```

PERFORM PROCRUSTES SUPERIMPOSITION AND REMOVE ASYMMETRY VARIATION
```{r}
ind <- dimnames(alldata)[[3]]
pet_sym <- bilat.symmetry(alldata, ind=ind, land.pairs = sym, object.sym=T, curves=slide)
```
PERFORM STANDARD GPA TO OBTAIN CRANIAL (CENTROID) SIZES
```{r}
gpa <- gpagen(alldata)
size <- gpa$Csize
```

TEST ALLOMETRY USING OLS MODEL
```{r}
allom <- procD.lm(pet_sym$symm.shape~log(size), iter=999); summary(allom)
allom_plot <- plot(allom, type = "regression", predictor = log(size), 
                   reg.type = "RegScore", pch = iconsPCA, bg=colsPCA, 
                   xlab = "log(centroid size)", ylab = "Regression score", cex=2)
```

USE 'SHAPE.PREDICTOR' TO VISUALLY EXAMINE PREDICTED SHAPE CHANGES THAT OCCUR WITH INCREASED CRANIAL SIZE
```{r}
preds <- shape.predictor(allom$GM$fitted, x = allom_plot$PredLine, 
                         predmin = min(allom_plot$PredLine), 
                         predmax = max(allom_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmax, preds$predmin, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)
```

PERFORM PRINCIPAL COMPONENT ANALYSIS AND PLOT PC1 VS PC2
```{r}
PCA <- gm.prcomp(pet_sym$symm.shape)
plot(PCA$x[,2]~PCA$x[,1], pch=iconsPCA, cex=size/200, axes.linetype=NA, bg=colsPCA, asp=1, xlab = paste("PC1", "(", (signif(PCA$d[1]/sum(PCA$d)*100, digits=4)), "%)", sep=""), ylab = paste("PC2", "(", (signif(PCA$d[2]/sum(PCA$d)*100, digits=3)), "%)", sep=""))
```

TEST FOR SIZE CORRELATIONS FOR MAJOR COMPONENTS OF SHAPE
```{r}
PC1cor <- lm(PCA$x[,1]~log(gpa$Csize)); print(summary(PC1cor)) # SIGNIFICANT
PC2cor <- lm(PCA$x[,2]~log(gpa$Csize)); print(summary(PC2cor)) # SIGNIFICANT
PC3cor <- lm(PCA$x[,3]~log(gpa$Csize)); print(summary(PC3cor)) # SIGNIFICANT
PC4cor <- lm(PCA$x[,4]~log(gpa$Csize)); print(summary(PC4cor)) # NOT SIGNIFICANT
```

PLOT HEATMAPS OF SHAPE DIFFERENCES ALONG PC1
```{r}
differences <- coordinates.difference(coordinates = PCA$shapes$shapes.comp1$max,
                                      reference = PCA$shapes$shapes.comp1$min, type = "spherical")
procrustes.var.plot(PCA$shapes$shapes.comp1$min, PCA$shapes$shapes.comp1$max, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)
```

PLOT HEATMAPS OF SHAPE DIFFERENCES ALONG PC2
```{r}
differences <- coordinates.difference(coordinates = PCA$shapes$shapes.comp2$max,
                                      reference = PCA$shapes$shapes.comp2$min, type = "spherical")
procrustes.var.plot(PCA$shapes$shapes.comp2$min, PCA$shapes$shapes.comp2$max, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)
```

PLOT HEATMAPS OF SHAPE DIFFERENCES ALONG PC3
```{r}
differences <- coordinates.difference(coordinates = PCA$shapes$shapes.comp3$max,
                                      reference = PCA$shapes$shapes.comp3$min, type = "spherical")
procrustes.var.plot(PCA$shapes$shapes.comp3$min, PCA$shapes$shapes.comp3$max, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)
```
