---
title: "Petrogale dental tests"
author: "D. Rex Mitchell"
date: "2023-05-04"
output: html_document
editor_options: 
  chunk_output_type: console
---
IMPORT PACKAGES
```{r}
library(geomorph)
library(FSA)
library(ggplot2)
```

LOAD DATA
```{r}
load("Petrogale_dental_data.rda")
```

TESTING
PREMOLAR TESTS

REMOVE SPECIMENS WITH NO PREMOLAR DATA AND GENERATE AN INDEX OF RELATIVE PREMOLAR SIZE
```{r}
n <- which(is.na(p3data))
size_p3 <- size[-n]
p3 <- p3data[-n]
sp <- species[-n]
p3rat <- p3/size_p3
```

PERFORM PAIRWISE COMPARISONS BETWEEN ALL SPECIES
```{r}
comp <- procD.lm(p3rat~sp, iter=999); summary(comp)
PW1 <- pairwise(comp, groups = sp)
summary(PW1)
```

SUMMARY STATISTICS OF ALL DATA
```{r}
Sum = Summarize(p3rat~sp)
Sum$se = Sum$sd / sqrt(Sum$n)
pd = position_dodge(1) 
```

PLOT SPECIES MEANS
```{r}
p <- ggplot(Sum, aes(x = factor(sp, level = c("P. penicillata","P. herberti","P. godmani","P. coenensis","P. inornata", "P. assimilis","P. sharmani","P. mareeba", "P. purpureicollis","P. lateralis", "P. rothschildi", "P. persephone", "P. xanthopus","P. brachyotis","P. wilkinsi","P. burbidgei","P. concinna")), y = mean)) +
  geom_bar(stat     = "identity",
           color    = "grey") +
  geom_errorbar(aes(ymin  = mean - se,
                    ymax  = mean + se),
                width = 0.2,
                linewidth  = 0.7,
                position = pd,
                color = "black"
  ) + ylim(0.0125, 0.021) + labs(y= "mean P3 size", x = ""); p
q <- p+theme(axis.text.x = element_text(angle = 90)); q
q + theme(axis.text.x = element_text(face = "italic")) +
  ylab("relative P3 size")
```



MOLAR TESTS

REMOVE SPECIMENS WITH NO MOLAR DATA AND GENERATE AN INDEX OF RELATIVE PREMOLAR SIZE
```{r}
m <- which(is.na(m2data))
size_m2 <- size[-m]
m2 <- m2data[-m]
sp <- species[-m]
m2rat <- m2/size_m2
```

PERFORM PAIRWISE COMPARISONS BETWEEN ALL SPECIES
```{r}
comp <- procD.lm(m2rat~sp, iter=999); summary(comp)
PW1 <- pairwise(comp, groups = sp)
summary(PW1)
```

SUMMARY STATISTICS OF ALL DATA
```{r}
Sum = Summarize(m2rat~sp)
Sum$se = Sum$sd / sqrt(Sum$n)
pd = position_dodge(1)
```

PLOT SPECIES MEANS
```{r}
p <- ggplot(Sum,                ### The data frame to use.
            aes(x = factor(sp, level = c("P. penicillata","P. herberti","P. godmani","P. coenensis","P. inornata", "P. assimilis","P. sharmani","P. mareeba", "P. purpureicollis","P. lateralis", "P. rothschildi", "P. persephone", "P. xanthopus","P. brachyotis","P. wilkinsi","P. burbidgei","P. concinna")),
                y     = mean)) +
  geom_bar(stat     = "identity",
           color    = "grey") +
  geom_errorbar(aes(ymin  = mean - se,
                    ymax  = mean + se),
                width = 0.2,
                linewidth  = 0.7,
                position = pd,
                color = "black"
  ) +ylim(0.0125, 0.0175) + labs(y= "mean M2 size", x = ""); p
q <- p+theme(axis.text.x = element_text(angle = 90)); q
q + theme(axis.text.x = element_text(face = "italic")) +
  ylab("relative M2 size")
```
